\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

That the flow generated by a Fokker-Planck PDE initial value problem can be recovered from a variational recursion of the form \eqref{ProxRecursionInfiniteDim} was first proposed in \cite{jordan1998variational}, showing that when the drift in the sample path dynamics is a gradient vector field and the diffusion is a scalar multiple of identity matrix, then ${\rm{dist}}(\cdot,\cdot)$ can be taken as the Wasserstein-2 metric arising in the theory of optimal transport \cite{villani2003topics} with $\Phi(\cdot)$ as the free energy functional. In particular, the functional $\Phi$ serves as a Lyapunov functional in the sense $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of the Fokker-Planck PDE initial value problem. This idea has since been generalized to many other types of PDE initial value problems, see e.g., \cite{ambrosio2008gradient,santambrogio2017euclidean}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-0.2in}
\section{Numerical Simulations}\label{sec:NumericalSimulations}
To illustrate the proposed computational framework, we next provide two numerical simulation case studies. In Sec. \ref{subsec:NumSimIEEEBus}, we consider the prediction of transient stochastic states for the IEEE 14 bus system for the nominal case as well as for the case when a line failure occurs. In Sec. \ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the 100 dimensional state space of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method. All simulations were performed in MATLAB R2019b on an iMac with 3.4 GHz Quad-Core Intel Core i5 processor and 8 GB memory.

@@

\vspace*{-0.2in}
\subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
We consider the Kron-reduced dynamics \eqref{ItoSDEcomponentlevel} for the IEEE 14 bus system shown in Fig. \ref{fig:Motivating}. In this case, we have $n=5$ nodes which correspond to the buses 1, 2, 3, 6 and 8 in Fig. \ref{fig:Motivating}. We obtained the parameters of the 14-bus system from MATPOWER \cite{zimmerman2010matpower}. The calculation of Kron-reduced admittance matrix and current vector in \eqref{DefParam} followed Sec.  \ref{subsec:samplepathdynamics}; see also \cite{dorfler2012kron}. The parameters $m_{i}, \gamma_{i}$ were obtained from the open-source Python-based
library ANDES \cite{cui2021andes}.

@@

Since \textbf{Case II} considers the failure of line 13, from Fig. \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4). The comparison of the $(\theta_4,\omega_4)$ bivariate marginals between \textbf{Cases I} and \textbf{II} shown in Fig. \ref{fig:IEEE14marginalsBivariateComparison} indeed confirms that the post contingency transient stochastic states for bus 6 have higher dispersion or spread of the probability mass.

@@

\bibitem{santambrogio2017euclidean}
F.~Santambrogio, ``$\{$Euclidean, Metric, and Wasserstein$\}$ Gradient Flows: An Overview", \emph{Bulletin of Mathematical Sciences}, Vol. 7, No. 1, pp. 87--154, 2017.

@@

\bibitem{krichene2020global}
W. Krichene, K.F.~Caluya, and A.~Halder, ``Global Convergence of Second-order Dynamics in Two-layer Neural Networks", \emph{arXiv preprint arXiv:2007.06852}, 2020.

>>>>

That the flow generated by a Fokker-Planck PDE initial value problem can be recovered from a variational recursion of the form \eqref{ProxRecursionInfiniteDim} was first proposed in \cite{jordan1998variational}, showing that when the drift in the sample path dynamics is a gradient vector field and the diffusion is a scalar multiple of identity matrix, then ${\rm{dist}}(\cdot,\cdot)$ can be taken as the Wasserstein-2 metric arising in the theory of optimal transport \cite{villani2003topics} with $\Phi(\cdot)$ as the free energy functional. In particular, the functional $\Phi$ serves as a Lyapunov functional in the sense $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of the Fokker-Planck PDE initial value problem. This idea has since been generalized to many other types of PDE initial value problems, see e.g., \cite{ambrosio2008gradient}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Simulations}\label{sec:NumericalSimulations}
To illustrate the proposed computational framework, we next provide two numerical simulation case studies. In Sec. \ref{subsec:NumSimIEEEBus}, we consider the prediction of transient stochastic states for the IEEE 14 bus system for the nominal case as well as for the case when a line failure occurs. In Sec. \ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the 100 dimensional state space of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method. All simulations were performed in MATLAB R2019b on an iMac with 3.4 GHz Quad-Core Intel Core i5 processor and 8 GB memory.

@@

\subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
We consider the Kron-reduced dynamics \eqref{ItoSDEcomponentlevel} for the IEEE 14 bus system shown in Fig. \ref{fig:Motivating}. In this case, we have $n=5$ nodes which correspond to the buses 1, 2, 3, 6 and 8 in Fig. \ref{fig:Motivating}. We obtained the parameters of the 14-bus system from MATPOWER \cite{zimmerman2010matpower}. The calculation of Kron-reduced admittance matrix and current vector in \eqref{DefParam} followed Sec.  \ref{subsec:samplepathdynamics}; see also \cite{dorfler2012kron}. The parameters $m_{i}, \gamma_{i}$ were obtained from the open-source Python-based
library ANDES \cite{cui2021andes}.

@@

Since \textbf{Case II} considers the failure of line 13, from Fig. \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4). The comparison of the $(\theta_4,\omega_4)$ bivariate marginals between \textbf{Cases I} and \textbf{II} shown in Fig. \ref{fig:IEEE14marginalsBivariateComparison} indeed confirms that the post-contingency transient stochastic states for bus 6 have higher dispersion or spread of the probability mass.

@@

%\bibitem{santambrogio2017euclidean}
%F.~Santambrogio, ``$\{$Euclidean, Metric, and Wasserstein$\}$ Gradient Flows: An Overview", \emph{Bulletin of Mathematical Sciences}, Vol. 7, No. 1, pp. 87--154, 2017.

@@

%\bibitem{krichene2020global}
%W. Krichene, K.F.~Caluya, and A.~Halder, ``Global Convergence of Second-order Dynamics in Two-layer Neural Networks", \emph{arXiv preprint arXiv:2007.06852}, 2020.
